---
title: 'Application of Neural Networks for Prediction of Authorization to Organ Donation '
author: "Pradeep Bagavan"
date: "December 21, 2016"
output:
  html_document:
    code_folding: show
    highlight: haddock
    theme: default
  pdf_document: default
  word_document: default
theme: journal
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Objective 
Predict whether the family of a organ donor will say Yes/No to Organ Donation
Methodology: Neural Network (Multilayer Perceptron)

In this project, no  hidden layers were used. Linear basis function and sigmoidal activation 
function were used in the MLP

### Specify the Input and Output Layer Size
```{r warning=FALSE, comment=FALSE, message=FALSE}
library(ggplot2)
input_layer_size <- 8 # Excluding bias 
output_layer_size <- 1
```

### Loading data

```{r warning=FALSE}
r <- read.csv("C:/Users/pbagavan/SSIE519/mlclass-ex4/data.csv")

# include bias unit
bias <- rep(1, times=nrow(r))
r <- cbind(bias,r)

# Standardize age (AG) and Admission to Referral Time (AR) 
r$AG <- ((r$AG*0.8)/95)+0.1
r$AR <- ((r$AR*0.08)/644.89)+0.1
r <- as.matrix(r)
```

### Data Slicing

```{r warning=FALSE}
set.seed(1050)
index <- sample(1:nrow(r),round(0.70*nrow(r)))
train <- r[index,]
test <- r[-index,]
```

### Training data

**Preview table**

```{r warning=FALSE}
head(round(train[1:5,],4))

# Storing the error in Epsilon_Store
Epsilon_Store <- matrix(data=0,nrow=nrow(train),ncol=output_layer_size) 
Overall_Error <- 0 


# Input and output variables
train_x <- train[,1:(input_layer_size+1)] 
train_Y <- train[,-(1:(input_layer_size+1))] 
 

# Initialize weights
set.seed(500) 
w1 <- matrix(data=round((4*runif(input_layer_size+1,0,1)-2)/3,4),
             nrow=output_layer_size,ncol=input_layer_size+1,
             byrow = T) 

#t=1{1st Epoch, first step in epoch: zero out the c1 matrix}
t=1 
for(t in 1:465){
c1 <- matrix(data=0,nrow=output_layer_size,ncol=input_layer_size+1,
             byrow = T)
 #i=1 {1st training example. First read in values for input layer activations}
  i <- 1
  for(i in 1:nrow(train)){
  # Forward propagate
  s <- train_x[i,] %*% t(w1)
  a <- 1/(1+exp(-s))
  
  # Backward propagate
  Epsilon <- a-train_Y[i]
  Epsilon_Store[i] <- Epsilon 
  d <- Epsilon*a*(1-a)
  # Update c matrix
  c1 <- c1+t(train_x[i,])*d[1,1]
  }

 #End of first epoch. Update weights
 w1 <- w1 - 0.01*c1
 Overall_Error[t] <- 0.5*sum(Epsilon_Store^2)/input_layer_size
}

# Input and out test variables
test_x <- test[,1:(input_layer_size+1)] 
test_Y <- test[,-(1:(input_layer_size+1))] 

# Save activations of output node 
test_a <- rep(0, times=nrow(test))
i <- 1
for(i in 1:nrow(test)){
  # Forward propagate
  s <- test_x[i,] %*% t(w1)
  test_a[i] <- 1/(1+exp(-s))
}

```

### Plotting activations
```{r warning=FALSE}
SampleNo <- 1:nrow(test_x)
OpActivations <- test_a
Authorization <- ifelse(OpActivations>=0.5,"darkgreen","red")
plot_data <- as.data.frame(cbind(SampleNo,round(OpActivations,2),Authorization))

#Overall_Error
ggplot() + geom_point(aes(x=1:t,y=Overall_Error)) + theme_bw() + 
  xlab("Number of iterations ")+ylab(expression(paste("Overall Error ",hat(E))))+
  annotate("text",x=440,y=5, label= "Learning Rate = 0.01")

ggplot(plot_data, aes(x=SampleNo,y=OpActivations)) + 
  geom_point(aes(group=Authorization,color=Authorization),shape=2) + 
  ylab("Output Node Activations")+
  geom_hline(aes(yintercept=0.5), lty=2,color="blue",lwd=1.2)+
    scale_color_manual(values = c("darkgreen", "red"), labels = c("Yes", "No"))+
  theme_bw()+
  theme(axis.title.x=element_blank(),axis.text.x=element_blank(),
        axis.ticks.x=element_blank())
```


# Converting Auth from continuous to categorical variable
```{r warning=FALSE}
plot_data$test_Y_cat <- ifelse(test_Y ==0.1,"No","Yes")
plot_data$test_Y_cat <- factor(plot_data$test_Y_cat, levels = c("Yes","No"))
# Converting Predicted varible into categorical variable
plot_data$pred_Y_cat <- ifelse(OpActivations>=0.5,"Yes","No")
plot_data$pred_Y_cat <- factor(plot_data$pred_Y_cat, levels = c("Yes","No"))
plot_data$category <- ifelse(plot_data$test_Y_cat=="Yes" & plot_data$pred_Y_cat=="Yes","A",
                      ifelse(plot_data$test_Y_cat=="No" & plot_data$pred_Y_cat=="Yes","B",
                      ifelse(plot_data$test_Y_cat=="Yes" & plot_data$pred_Y_cat=="No","C","D")))

#Classification Accuracy
class.accuracy <- round((sum(plot_data$category=="A"|plot_data$category=="D")/nrow(plot_data))*100,2)
class.accuracy
# Misclassification Rate
misclass <- round((sum(plot_data$category=="B"|plot_data$category=="C")/nrow(plot_data))*100,2)
misclass
```


### Threshold Distribution
```{r warning=FALSE}
df <- as.data.frame(cbind(test_a,test_Y))

plot_distribution <- function(df,threshold){
  v <- rep(NA, nrow(df))
  v <- ifelse(df$test_a >= threshold & df$test_Y == 0.9, "TP", v)
  v <- ifelse(df$test_a >= threshold & df$test_Y == 0.1, "FP", v)
  v <- ifelse(df$test_a < threshold & df$test_Y == 0.9, "FN", v)
  v <- ifelse(df$test_a < threshold & df$test_Y == 0.1, "TN", v)
  
  df$pred_type <- v
  
  ggplot(df, aes(x=test_Y, y=test_a))+
  geom_violin(fill=rgb(1,1,1,alpha = 0.6), color=NA)+
  geom_jitter(aes(color=pred_type),alpha=0.6)+
  geom_hline(yintercept=threshold,color="red",alpha=0.6)+
  scale_color_discrete(name=" ")+
  labs(title=sprintf("Threshold = %.2f", threshold))+
  xlab("Actual Value")+ylab("Predicted Value")
}
```

### ROC Calculation
```{r warning=FALSE}
calculate_roc <- function(df, cost_of_fp, cost_of_fn, n=100) {
  tpr <- function(df, threshold) {
    sum(df$test_a >= threshold & df$test_Y == 0.9) / sum(df$test_Y == 0.9)
  }
  
  fpr <- function(df, threshold) {
    sum(df$test_a >= threshold & df$test_Y == 0.1) / sum(df$test_Y == 0.1)
  }
  
  cost <- function(df, threshold, cost_of_fp, cost_of_fn) {
    sum(df$test_a >= threshold & df$test_Y == 0.1) * cost_of_fp + 
      sum(df$test_a < threshold & df$test_Y == 0.9) * cost_of_fn
  }
  
  roc <- data.frame(threshold = seq(0,1,length.out=n), tpr=NA, fpr=NA)
  roc$tpr <- sapply(roc$threshold, function(th) tpr(df, th))
  roc$fpr <- sapply(roc$threshold, function(th) fpr(df, th))
  roc$cost <- sapply(roc$threshold, function(th) cost(df, th, cost_of_fp, cost_of_fn))
  
  return(roc)
}

# ROC plotting
plot_roc <- function(roc, threshold, cost_of_fp, cost_of_fn) {
  library(gridExtra)
  library(grid)
  norm_vec <- function(v) (v - min(v))/diff(range(v))
  
  idx_threshold = which.min(abs(roc$threshold-threshold))
  
  col_ramp <- colorRampPalette(c("green","orange","red","black"))(100)
  col_by_cost <- col_ramp[ceiling(norm_vec(roc$cost)*99)+1]
  p_roc <- ggplot(roc, aes(fpr,tpr)) + 
    geom_line(color=rgb(0,0,1,alpha=0.3)) +
    geom_point(color=col_by_cost, size=4, alpha=0.5) +
    coord_fixed() +
    geom_line(aes(threshold,threshold), color=rgb(0,0,1,alpha=0.5)) +
    labs(title = sprintf(" ")) + xlab("1-Specificity") + ylab("Sensitivity") +
    geom_hline(yintercept=roc[idx_threshold,"tpr"], alpha=0.5, linetype="dashed") +
    geom_vline(xintercept=roc[idx_threshold,"fpr"], alpha=0.5, linetype="dashed")
  
  p_cost <- ggplot(roc, aes(threshold, cost)) +
    geom_line(color=rgb(0,0,1,alpha=0.3)) +
    geom_point(color=col_by_cost, size=4, alpha=0.5) +
    labs(title = sprintf(" ")) +
    geom_vline(xintercept=threshold, alpha=0.5, linetype="dashed")+
    xlab("Threshold")+ylab("Cost")
  sub_title <- sprintf("      threshold at %.2f - cost of FP = %d, cost of FN = %d", threshold, cost_of_fp, cost_of_fn)
  
  grid.arrange(p_roc, p_cost, ncol=2, sub=textGrob(sub_title, gp=gpar(cex=1), just="bottom"))
}
```
### Cost-Specific Threshold Values

False Positives (FP): The model predicting incorrectly that the family will authorize when it actually does not

If I assume cost of 2 for FP and 1 for False Negatives (FN)

```{r warning=FALSE}
roc <- calculate_roc(df, 2,1,n=100)
Best_Threshold <- roc$threshold[which.min(roc$cost)]
Best_Threshold
plot_roc(roc, Best_Threshold, 2, 1)
plot_distribution(df, Best_Threshold)
```


